\bf An algebraic formulation of dependent type theory

### 2014-11-05 Presentation/presentation.tex
nearly finished presentation

\documentclass[handout]{beamer}
\setbeamercolor{frametitle}{fg=teal!80!black}
\setbeamercolor{item}{fg=red!90!black}
\setbeamercolor{title}{fg=teal!80!black}
\setbeamercolor{block title}{fg=teal!80!black}

@@

\newcommand\goal[1]{\begin{huge}\begin{center}\textbf{Goal: {\color{teal}#1}}\end{center}\end{huge}}
\newcommand\important[1]{\textbf{\color{red!90!black}#1}}

@@

\begin{frame}
\frametitle{\bf The basic judgments}
\begin{align*}
\jalign\jctx{\Gamma} 
& \jalign\jctxeq{\Gamma}{\Gamma'}
  \\
\jalign\jfam{\Gamma}{A} 
& \jalign\jfameq{\Gamma}{A}{B}
  \\
\jalign\jterm{\Gamma}{A}{x} 
& \jalign\jtermeq{\Gamma}{A}{x}{y}.
\end{align*}
\\[\baselineskip]
There are rules expressing that all three kinds of judgmental equality are
convertible equivalence relations.

@@

\begin{comment}
\begin{frame}
\frametitle{\bf Rules: judgmental equality is an equivalence relation}
\begin{footnotesize}
\begin{align*}
& \inference
  { \jctx{\Gamma}
    }
  { \jctxeq{\Gamma}{\Gamma}
    } 
& & \inference
    { \jctxeq{\Gamma}{\Delta}
      }
    { \jctxeq{\Delta}{\Gamma}
      } 
& & \inference
    { \jctxeq{\Gamma}{\Delta}
      \jctxeq{\Delta}{\greek{E}}
      }
    { \jctxeq{\Gamma}{\greek{E}}
      }
    \\
& \inference
  { \jfam{\Gamma}{A}
    }
  { \jfameq{\Gamma}{A}{A}
    } 
& & \inference
    { \jfameq{\Gamma}{A}{B}
      }
    { \jfameq{\Gamma}{B}{A}
      }
& & \inference
    { \jfameq{\Gamma}{A}{B}
      \jfameq{\Gamma}{B}{C}
      }
    { \jfameq{\Gamma}{A}{C}
      }
    \\
& \inference
  { \jterm{\Gamma}{A}{x}
    }
  { \jtermeq{\Gamma}{A}{x}{x}
    }
& & \inference
    { \jtermeq{\Gamma}{A}{x}{y}
      }
    { \jtermeq{\Gamma}{A}{y}{x}
      }
& & \inference
    { \jtermeq{\Gamma}{A}{x}{y}
      \jtermeq{\Gamma}{A}{y}{z}
      }
    { \jtermeq{\Gamma}{A}{x}{z}
      }
\end{align*}
\begin{align*}
& \inference
  { \jctxeq{\Gamma}{\Delta}
    \jfam{\Gamma}{A}
    }
  { \jfam{\Delta}{A}
    }
& & \inference
    { \jctxeq{\Gamma}{\Delta}
      \jfameq{\Gamma}{A}{B}
      }
    { \jfameq{\Delta}{A}{B}
      }
    \\
& \inference
  { \jctxeq{\Gamma}{\Delta}
    \jterm{\Gamma}{A}{x}
    }
  { \jterm{\Delta}{A}{x}
    }
& & \inference
    { \jctxeq{\Gamma}{\Delta}
      \jtermeq{\Gamma}{A}{x}{y}
      }
    { \jtermeq{\Delta}{A}{x}{y}
      }
    \\
& \inference
  { \jfameq{\Gamma}{A}{B}
    \jterm{\Gamma}{A}{x}
    }
  { \jterm{\Gamma}{B}{x}
    }
& & \inference
    { \jfameq{\Gamma}{A}{B}
      \jtermeq{\Gamma}{A}{x}{y}
      }
    { \jtermeq{\Gamma}{B}{x}{y}
      }
\end{align*}
\end{footnotesize}
\end{frame}
\end{comment}

@@

\begin{frame}
\frametitle{\bf Current goal}
\goal{Define extension algebras}
\end{frame}

@@

\begin{frame}
\frametitle{\bf Notation for pre-extension algebras}
We introduce the following notation:
\begin{align*}
\stesysf_2 
  & := \stesysff
  \\
\eft[2] 
  & := \pullbackpr{1}{\ectxext}{\eft} : \stesysf_2\to\stesysf
  \\
\stesysf_3 & := \pullback{\stesysf_2}{\stesysf_2}{\efamext}{\eft[2]}
  \\
\eft[3]
  & := \pullbackpr{1}{\efamext}{\eft[2]} : \stesysf_3\to\stesysf_2.
\end{align*}
Then it follows that the outer square in the diagram
\begin{equation*}
\begin{tikzcd}[ampersand replacement=\&,column sep=large]
\stesysf_3
  \ar[dotted]{dr}{\eext{2}}
  \ar{rr}{\pullback{\pullbackpr{2}{\ectxext}{\eft}}{\pullbackpr{2}{\ectxext}{\eft}}{\ectxext}{\eft}}
  \ar{dd}[swap]{\eft[3]}
  \& 
  \&
\stesysf_2
  \ar{d}{\efamext}
  \\
  \&
\stesysf_2
  \ar{d}[swap]{\eft[2]}
  \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
  \&
\stesysf
  \ar{d}{\eft}
  \\
\stesysf_2
  \ar{r}[swap]{\eft[2]}
  \&
\stesysf
  \ar{r}[swap]{\ectxext}
  \&
\stesysc
\end{tikzcd}
\end{equation*}
commutes. We define $\eext{2}$ to be the unique morphism rendering the above diagram
commutative. 
\end{frame}

@@

\begin{frame}
\frametitle{\bf Current goal}
\goal{Develop properties of extension algebras}

We need to know:
\begin{itemize}
\item What homomorphisms of (pre-)extension algebras are
\item That 
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{\bf Slicing of pre-extension algebras}
Suppose that $\stesys$ is a pre-extension algebra. Then we define
the pre-extension algebra $\famesys{\stesys}$ to consist of the
fundamental structure
\begin{equation*}
\begin{tikzcd}[ampersand replacement=\&]
\stesyst_2
  \ar{d}{\ebd[2]}
  \\
\stesysf_2
  \ar{d}{\eft[2]}
  \\
\stesysf
\end{tikzcd}
\end{equation*}
where
\begin{align*}
\stesyst_2 
  & := \pullback{\stesysf}{\stesyst}{\ectxext}{\eft\circ\ebd},
  &
\ebd[2]
  & := \ectxext^\ast(\ebd),
\end{align*}
with the extension operations
\begin{align*}
\efamext 
  & 
: \stesysf_2\to\stesysf,
  &
\eext{2} 
  & 
: \stesysf_3\to\stesysf_2.
\end{align*}
\end{frame}

\begin{frame}
\frametitle{\bf Local extension algebras}
\begin{theorem}
If $\stesys$ is an extension algebra, then so is $\famesys{\stesys}$.
\end{theorem}
\end{frame}

\begin{frame}
\frametitle{\bf Pre-extension homomorphisms}
\begin{footnotesize}
Let $\stesys$ and $\stesys'$ be pre-extension algebras. A \emph{pre-extension 
homomorphism $f$ from $\stesys'$ to $\stesys$} is a triple $(f_0,f_1,f^t)$ 
consisting of morphisms
\begin{equation*}
\begin{tikzcd}[ampersand replacement=\&]
\stesyst' 
  \ar{r}{f^t}
  \ar{d}[swap]{\ebd'}
  \&
\stesyst
  \ar{d}{\ebd}
  \\
\stesysf'
  \ar{r}{f_1}
  \ar{d}[swap]{\eft'}
  \&
\stesysf
  \ar{d}{\eft}
  \\
\stesysc' 
  \ar{r}[swap]{f_0}
  \&
\stesysc
\end{tikzcd}
\end{equation*}
such that the indicated squares commute, for which furthermore the squares
\begin{equation*}
\begin{tikzcd}[ampersand replacement=\&]
\stesysf' 
  \ar{r}{f_1}
  \ar{d}[swap]{\ectxext'}
  \&
\stesysf
  \ar{d}{\ectxext}
  \\
\stesysc'
  \ar{r}[swap]{f_0}
  \&
\stesysc
\end{tikzcd}
\quad
\text{and}
\quad
\begin{tikzcd}[ampersand replacement=\&,column sep=huge]
\stesysf_2'
  \ar{r}{f_1\times_{\ectxext,\eft} f_1}
  \ar{d}[swap]{\efamext'}
  \&
\stesysf_2
  \ar{d}{\efamext}
  \\
\stesysf'
  \ar{r}[swap]{f_1}
  \&
\stesysf
\end{tikzcd}
\end{equation*}
commute. Composition and the identity homomorphism are defined in the expected way.
\end{footnotesize}
\end{frame}

\begin{frame}
\begin{definition}
A pre-extension homomorphism between extension algebras is called an extension
homomorphism.
\end{definition}

@@

\begin{frame}
\frametitle{\bf Change of base of pre-extension algebras}
Let $\stesys$ be a pre-extension algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:g$.
Then there is a pre-extension algebra $\cobesys{Y}{\stesys}{g}{p}$ with projections, such that for every diagram
\begin{equation*}
\begin{tikzcd}[ampersand replacement=\&,column sep=large]
\stesys'
  \ar[bend right=10]{ddr}[swap]{p'}
  \ar[bend left=10]{rrd}{f}
  \ar[dotted]{dr}[near end]{[p',f]}
  \\
  {}\&
\cobesys{Y}{\stesys}{g}{p}
  \ar{d}{\pullbackpr{1}{g}{p}}
  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
  \&
\stesys
  \ar{d}{p}
  \\
  {}\&
Y \ar{r}[swap]{g}
  \&
X
\end{tikzcd}
\end{equation*}
of which the outer square commutes, the pre-extension homomorphism $[p',f]$ exists
and is unique with the property that it renders the diagram commutative.
\end{frame}

\begin{frame}
\frametitle{\bf Change of base of extension algebras}
\begin{theorem}
The change of base of an extension algebra is again an extension algebra
\end{theorem}

@@

\begin{frame}
\frametitle{\bf Introduction to the theory of weakening}
\begin{itemize}
\item We will introduce an operation for weakening which acts at the three
levels of fundamental structures (contexts, families and terms).
  \pause
\item It will preserve extension, so it will be implemented as an extension
homomorphism.
  \pause
\item It will preserve itself, so weakening will also be a weakening
homomorphism.
  \pause
\item There will be a notion of `Currying for weakening', which explains
what happens when something is weakened by an extended object.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{\bf Rules for the theory of weakening}
\begin{footnotesize}
The introduction rules for weakening:
\begin{align*}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{\Gamma}{B}
    }
  { \jfam{{\Gamma}{A}}{\ctxwk{A}{B}}
    }
& & \inference
    { \jfameq{\Gamma}{A}{A'}
      \jfameq{\Gamma}{B}{B'}
      }
    { \jfameq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}
      }
    \\
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{\Gamma}{B}}{Q}
    }
  { \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}
    }
& & \inference
    { \jfameq{\Gamma}{A}{A'}
      \jfameq{{\Gamma}{B}}{Q}{Q'}
      }
    { \jfameq
        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
        {\ctxwk{A}{Q}}
        {\ctxwk{A'}{Q'}}
      }
    \\
& \inference
  { \jfam{\Gamma}{A}
    \jterm{{\Gamma}{B}}{Q}{g}
    }
  { \jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A}{g}}
    }
& & \inference
    { \jfameq{\Gamma}{A}{A'}
      \jtermeq{{\Gamma}{B}}{Q}{g}{g'}
      }
    { \jtermeq
        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
        {\ctxwk{A}{Q}}
        {\ctxwk{A}{g}}
        {\ctxwk{A'}{g'}}
      }
\end{align*}
\pause
Weakening preserves extension:
\begin{align*}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{\Gamma}{B}}{Q}
    }
  { \jfameq
      {\ctxext{\Gamma}{A}}
      {\ctxwk{A}{\ctxext{B}{Q}}}
      {\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
    }
  \\
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{{\Gamma}{B}}{Q}}{R}
    }
  { \jfameq
      {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
      {\ctxwk{A}{\ctxext{Q}{R}}}
      {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
    }
\end{align*}
\end{footnotesize}
\end{frame}

\begin{frame}
\frametitle{\bf Pre-weakening}
Let $\stesys$ be an extension algebra. A pre-weakening operation
on $\stesys$ is an extension homomorphism 
\begin{equation*}
\mathbf{w}(\stesys)
    :
  \cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
    \to
  \famesys{\famesys{\stesys}}
\end{equation*}
for which the diagram
\begin{equation*}
\begin{tikzcd}[ampersand replacement=\&,column sep=large]
\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
  \ar{r}{\mathbf{w}(\stesys)}
  \ar{dr}[swap]{\pullbackpr{1}{\eft}{\eft}}
  \&
\famesys{\famesys{\stesys}}
  \ar{d}{\eft[2]}
  \\
  {}\& 
\stesysf
\end{tikzcd}
\end{equation*}
commutes.
\end{frame}

\begin{frame}
\frametitle{\bf Weakening for the families algebra}
Let $\stesys$ be an extension algebra with pre-weakening operation
$\mathbf{w}(\stesys)$. Then $\famesys{\stesys}$ has the pre-weakening operation
$\mathbf{w}(\famesys{\stesys})$ which is uniquely determined by rendering the
diagram
\begin{small}
\begin{equation*}
\begin{tikzcd}[ampersand replacement=\&,column sep=large]
\cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[2]}{\eft[2]}
  \ar{rr}{%
      \pullback{\beta_1}{\boldsymbol{\beta}}{\eft}{\eft}
    }
  \ar[bend right]{ddr}[swap]{\pullbackpr{1}{\eft[2]}{\eft[2]}}
  \ar[dotted]{dr}{\mathbf{w}(\famesys{\stesys})}
  \&
  {}\&
\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
  \ar{r}{\mathbf{w}(\stesys)}
  \&
\famesys{\famesys{\stesys}}
  \ar{d}{\boldsymbol{\beta}}
  \\
  {}\&
\famesys{\famesys{\famesys{\stesys}}}
  \ar{r}{\boldsymbol{\beta}_\mathbf{2}}
  \ar{d}[swap]{\eft[3]}
  \&
\famesys{\famesys{\stesys}}
  \ar{d}{\eft[2]}
  \ar{r}{\boldsymbol{\beta}}
  \&
\famesys{\stesys}
  \ar{d}{\eft}
  \\
  {}\&
\stesysf_2
  \ar{r}[swap]{\efamext}
  \&
\stesysf
  \ar{r}[swap]{\ectxext}
  \&
\stesysc
\end{tikzcd}
\end{equation*}
\end{small}
commutative.
\end{frame}

\begin{frame}
\frametitle{\bf Rules for weakening: Currying}
\begin{align*}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{\Gamma}{A}}{P}
    \jfam{\Gamma}{B}
    }
  { \jfameq
      {{{\Gamma}{A}}{P}}
      {\ctxwk{\ctxext{A}{P}}{B}}
      {\ctxwk{P}{{A}{B}}}
    }
\\
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{\Gamma}{A}}{P}
    \jfam{{\Gamma}{B}}{Q}
    }
  { \jfameq
      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
      {\ctxwk{\ctxext{A}{P}}{Q}}
      {\ctxwk{P}{{A}{Q}}}
    }
\\
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{\Gamma}{A}}{P}
    \jterm{{\Gamma}{B}}{Q}{g}
    }
  { \jtermeq
      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
      {\ctxwk{P}{{A}{Q}}}
      {\ctxwk{\ctxext{A}{P}}{g}}
      {\ctxwk{P}{{A}{g}}}
    } 
\end{align*}
\end{frame}

>>>>

\documentclass[handout]{beamer}
\setbeamercolor{frametitle}{fg=teal!90!black}
\setbeamercolor{item}{fg=red!90!black}
\setbeamercolor{title}{fg=teal!90!black}
\setbeamercolor{block title}{fg=teal!90!black}

@@

\newcommand\subgoal[1]{\begin{huge}\begin{center}\textbf{Subgoal: {\color{teal!90!black}#1}}\end{center}\end{huge}}
\newcommand\important[1]{\textbf{\color{red!90!black}#1}}

@@

\begin{frame}
\frametitle{\bf The basic judgments}
\begin{align*}
\jalign\jctx{\Gamma} 
& \jalign\jctxeq{\Gamma}{\Gamma'}
  \\
\jalign\jfam{\Gamma}{A} 
& \jalign\jfameq{\Gamma}{A}{B}
  \\
\jalign\jterm{\Gamma}{A}{a} 
& \jalign\jtermeq{\Gamma}{A}{a}{y}.
\end{align*}
\\[\baselineskip]
There are rules expressing that all three kinds of judgmental equality are
convertible equivalence relations.

@@

\begin{comment}
\begin{frame}
\frametitle{\bf Rules: judgmental equality is an equivalence relation}
\begin{footnotesize}
\begin{align*}
& \inference
  { \jctx{\Gamma}
    }
  { \jctxeq{\Gamma}{\Gamma}
    } 
& & \inference
    { \jctxeq{\Gamma}{\Delta}
      }
    { \jctxeq{\Delta}{\Gamma}
      } 
& & \inference
    { \jctxeq{\Gamma}{\Delta}
      \jctxeq{\Delta}{\greek{E}}
      }
    { \jctxeq{\Gamma}{\greek{E}}
      }
    \\
& \inference
  { \jfam{\Gamma}{A}
    }
  { \jfameq{\Gamma}{A}{A}
    } 
& & \inference
    { \jfameq{\Gamma}{A}{B}
      }
    { \jfameq{\Gamma}{B}{A}
      }
& & \inference
    { \jfameq{\Gamma}{A}{B}
      \jfameq{\Gamma}{B}{C}
      }
    { \jfameq{\Gamma}{A}{C}
      }
    \\
& \inference
  { \jterm{\Gamma}{A}{a}
    }
  { \jtermeq{\Gamma}{A}{a}{a}
    }
& & \inference
    { \jtermeq{\Gamma}{A}{a}{y}
      }
    { \jtermeq{\Gamma}{A}{y}{a}
      }
& & \inference
    { \jtermeq{\Gamma}{A}{a}{y}
      \jtermeq{\Gamma}{A}{y}{z}
      }
    { \jtermeq{\Gamma}{A}{a}{z}
      }
\end{align*}
\begin{align*}
& \inference
  { \jctxeq{\Gamma}{\Delta}
    \jfam{\Gamma}{A}
    }
  { \jfam{\Delta}{A}
    }
& & \inference
    { \jctxeq{\Gamma}{\Delta}
      \jfameq{\Gamma}{A}{B}
      }
    { \jfameq{\Delta}{A}{B}
      }
    \\
& \inference
  { \jctxeq{\Gamma}{\Delta}
    \jterm{\Gamma}{A}{a}
    }
  { \jterm{\Delta}{A}{a}
    }
& & \inference
    { \jctxeq{\Gamma}{\Delta}
      \jtermeq{\Gamma}{A}{a}{y}
      }
    { \jtermeq{\Delta}{A}{a}{y}
      }
    \\
& \inference
  { \jfameq{\Gamma}{A}{B}
    \jterm{\Gamma}{A}{a}
    }
  { \jterm{\Gamma}{B}{a}
    }
& & \inference
    { \jfameq{\Gamma}{A}{B}
      \jtermeq{\Gamma}{A}{a}{y}
      }
    { \jtermeq{\Gamma}{B}{a}{y}
      }
\end{align*}
\end{footnotesize}
\end{frame}
\end{comment}

@@

@@

\begin{frame}
\frametitle{\bf Notation for pre-extension algebras}
We introduce the following notation:
\begin{align*}
\stesysf_2 
  & := \stesysff
  \pause
  \\
\eft[2] 
  & := \pullbackpr{1}{\ectxext}{\eft} : \stesysf_2\to\stesysf
  \pause
  \\
\stesysf_3 & := \pullback{\stesysf_2}{\stesysf_2}{\efamext}{\eft[2]}
  \pause
  \\
\eft[3]
  & := \pullbackpr{1}{\efamext}{\eft[2]} : \stesysf_3\to\stesysf_2.
\end{align*}
\pause
Then it follows that the outer square in the diagram
\begin{equation*}
\begin{tikzcd}[ampersand replacement=\&,column sep=large]
\stesysf_3
  \ar[dotted]{dr}{\eext{2}}
  \ar{rr}{\pullback{\pullbackpr{2}{\ectxext}{\eft}}{\pullbackpr{2}{\ectxext}{\eft}}{\ectxext}{\eft}}
  \ar{dd}[swap]{\eft[3]}
  \& 
  \&
\stesysf_2
  \ar{d}{\efamext}
  \\
  \&
\stesysf_2
  \ar{d}[swap]{\eft[2]}
  \ar{r}{\pullbackpr{2}{\ectxext}{\eft}}
  \&
\stesysf
  \ar{d}{\eft}
  \\
\stesysf_2
  \ar{r}[swap]{\eft[2]}
  \&
\stesysf
  \ar{r}[swap]{\ectxext}
  \&
\stesysc
\end{tikzcd}
\end{equation*}
commutes. \pause
We define \important{$\eext{2}$} to be the unique morphism rendering the above diagram
commutative. 
\end{frame}

@@

\begin{frame}
\frametitle{Subgoal: develop properties of extension algebras}
We need to know:
\begin{itemize}
\item What homomorphisms of (pre-)extension algebras are.
\pause
\item That each extension algebra gives locally an extension algebra (stable under slicing).
\pause
\item That the change of base of an extension is an extension algebra. Change of base allows
for \important{`parametrized extension homomorphisms'} of which weakening and substitution
are going to be examples.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{\bf Pre-extension homomorphisms}
\begin{footnotesize}
Let $\stesys$ and $\stesys'$ be pre-extension algebras. A \important{pre-extension 
homomorphism $f$ from $\stesys'$ to $\stesys$} is a triple $(f_0,f_1,f^t)$ 
consisting of morphisms
\begin{equation*}
\begin{tikzcd}[ampersand replacement=\&]
\stesyst' 
  \ar{r}{f^t}
  \ar{d}[swap]{\ebd'}
  \&
\stesyst
  \ar{d}{\ebd}
  \\
\stesysf'
  \ar{r}{f_1}
  \ar{d}[swap]{\eft'}
  \&
\stesysf
  \ar{d}{\eft}
  \\
\stesysc' 
  \ar{r}[swap]{f_0}
  \&
\stesysc
\end{tikzcd}
\end{equation*}
such that the indicated squares commute, for which furthermore the squares
\begin{equation*}
\begin{tikzcd}[ampersand replacement=\&]
\stesysf' 
  \ar{r}{f_1}
  \ar{d}[swap]{\ectxext'}
  \&
\stesysf
  \ar{d}{\ectxext}
  \\
\stesysc'
  \ar{r}[swap]{f_0}
  \&
\stesysc
\end{tikzcd}
\quad
\text{and}
\quad
\begin{tikzcd}[ampersand replacement=\&,column sep=huge]
\stesysf_2'
  \ar{r}{f_1\times_{\ectxext,\eft} f_1}
  \ar{d}[swap]{\efamext'}
  \&
\stesysf_2
  \ar{d}{\efamext}
  \\
\stesysf'
  \ar{r}[swap]{f_1}
  \&
\stesysf
\end{tikzcd}
\end{equation*}
commute.
\end{footnotesize}
\end{frame}

\begin{frame}
\frametitle{Extension homomorphisms}
\begin{definition}
A pre-extension homomorphism between extension algebras is called an \important{extension
homomorphism}.
\end{definition}
\end{frame}

\begin{frame}
\frametitle{\bf Slicing of pre-extension algebras}
Suppose that $\stesys$ is a pre-extension algebra. Then we define
the pre-extension algebra \important{$\famesys{\stesys}$} to consist of the
fundamental structure
\begin{equation*}
\begin{tikzcd}[ampersand replacement=\&]
\stesyst_2
  \ar{d}{\ebd[2]}
  \\
\stesysf_2
  \ar{d}{\eft[2]}
  \\
\stesysf
\end{tikzcd}
\end{equation*}
where
\begin{align*}
\stesyst_2 
  & := \pullback{\stesysf}{\stesyst}{\ectxext}{\eft\circ\ebd},
  &
\ebd[2]
  & := \ectxext^\ast(\ebd),
\end{align*}
with the extension operations
\begin{align*}
\efamext 
  & 
: \stesysf_2\to\stesysf,
  &
\eext{2} 
  & 
: \stesysf_3\to\stesysf_2.
\end{align*}
\end{frame}

\begin{frame}
\frametitle{\bf Local extension algebras}
\begin{theorem}
If $\stesys$ is an extension algebra, then so is $\famesys{\stesys}$.
\end{theorem}

@@

\begin{frame}
\frametitle{\bf Change of base of pre-extension algebras}
Let $\stesys$ be a pre-extension algebra and consider $p:\stesysc\rightarrow X\leftarrow Y:g$.
Then there is a pre-extension algebra \important{$\cobesys{Y}{\stesys}{g}{p}$} with projections, such that for every diagram
\begin{equation*}
\begin{tikzcd}[ampersand replacement=\&,column sep=large]
\stesys'
  \ar[bend right=10]{ddr}[swap]{p'}
  \ar[bend left=10]{rrd}{f}
  \ar[dotted]{dr}[near end]{[p',f]}
  \\
  {}\&
\cobesys{Y}{\stesys}{g}{p}
  \ar{d}{\pullbackpr{1}{g}{p}}
  \ar{r}[swap]{\pullbackpr{2}{g}{p}}
  \&
\stesys
  \ar{d}{p}
  \\
  {}\&
Y \ar{r}[swap]{g}
  \&
X
\end{tikzcd}
\end{equation*}
of which the outer square commutes, the pre-extension homomorphism \important{$[p',f]$} exists
and is unique with the property that it renders the diagram commutative.
\end{frame}

\begin{frame}
\frametitle{\bf Change of base of extension algebras}
\begin{theorem}
The change of base of an extension algebra is again an extension algebra.
\end{theorem}

@@

\begin{frame}
\frametitle{\bf Overview of the theory of weakening}
\begin{itemize}
\item In the syntax, we will introduce an operation for weakening which acts at the three
levels of fundamental structures (contexts, families and terms). \pause
\important{When $B$ is weakened by $A$, we denote this by $\ctxwk{A}{B}$.}
  \pause
\item Weakening will preserve extension, so it will be implemented as an extension
homomorphism.
  \pause
\item There will be a notion of \important{`Currying for weakening'}, which explains
what happens when something is weakened by an extended object.
  \pause
\item Pre-weakening algebras will be extension algebras with a weakening operation
satisfying an implementation of the currying condition. \pause Pre-weakening homomorphisms
will preserve this structure.
  \pause
\item Weakening will preserve itself, so weakening we will require that weakening is
a pre-weakening homomorphism.
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{\bf Rules for the theory of weakening}
\begin{footnotesize}
The introduction rules for weakening:
\begin{align*}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{\Gamma}{B}
    }
  { \jfam{{\Gamma}{A}}{\ctxwk{A}{B}}
    }
& & \inference
    { \jfameq{\Gamma}{A}{A'}
      \jfameq{\Gamma}{B}{B'}
      }
    { \jfameq{{\Gamma}{A}}{\ctxwk{A}{B}}{\ctxwk{A'}{B'}}
      }
  \pause
  \\
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{\Gamma}{B}}{Q}
    }
  { \jfam{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}
    }
& & \inference
    { \jfameq{\Gamma}{A}{A'}
      \jfameq{{\Gamma}{B}}{Q}{Q'}
      }
    { \jfameq
        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
        {\ctxwk{A}{Q}}
        {\ctxwk{A'}{Q'}}
      }
    \\
& \inference
  { \jfam{\Gamma}{A}
    \jterm{{\Gamma}{B}}{Q}{g}
    }
  { \jterm{{{\Gamma}{A}}{\ctxwk{A}{B}}}{\ctxwk{A}{Q}}{\ctxwk{A}{g}}
    }
& & \inference
    { \jfameq{\Gamma}{A}{A'}
      \jtermeq{{\Gamma}{B}}{Q}{g}{g'}
      }
    { \jtermeq
        {{{\Gamma}{A}}{\ctxwk{A}{B}}}
        {\ctxwk{A}{Q}}
        {\ctxwk{A}{g}}
        {\ctxwk{A'}{g'}}
      }
\end{align*}
\pause
Weakening preserves extension:
\begin{align*}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{\Gamma}{B}}{Q}
    }
  { \jfameq
      {\ctxext{\Gamma}{A}}
      {\ctxwk{A}{\ctxext{B}{Q}}}
      {\ctxext{\ctxwk{A}{B}}{\ctxwk{A}{Q}}}
    }
  \\
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{{\Gamma}{B}}{Q}}{R}
    }
  { \jfameq
      {\ctxext{{\Gamma}{A}}{\ctxwk{A}{B}}}
      {\ctxwk{A}{\ctxext{Q}{R}}}
      {\ctxext{\ctxwk{A}{Q}}{\ctxwk{A}{R}}}
    }
\end{align*}
\end{footnotesize}
\end{frame}

\begin{frame}
\frametitle{\bf The weakening operation}
Let $\stesys$ be an extension algebra. A \important{weakening operation
on $\stesys$} is an extension homomorphism 
\begin{equation*}
\mathbf{w}(\stesys)
    :
  \cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
    \to
  \famesys{\famesys{\stesys}}
\end{equation*}
for which the diagram
\begin{equation*}
\begin{tikzcd}[ampersand replacement=\&,column sep=large]
\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
  \ar{r}{\mathbf{w}(\stesys)}
  \ar{dr}[swap]{\pullbackpr{1}{\eft}{\eft}}
  \&
\famesys{\famesys{\stesys}}
  \ar{d}{\eft[2]}
  \\
  {}\& 
\stesysf
\end{tikzcd}
\end{equation*}
commutes.
\end{frame}

\begin{frame}
\frametitle{\bf Rules for weakening: Currying}
\begin{align*}
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{\Gamma}{A}}{P}
    \jfam{\Gamma}{B}
    }
  { \jfameq
      {{{\Gamma}{A}}{P}}
      {\ctxwk{\ctxext{A}{P}}{B}}
      {\ctxwk{P}{{A}{B}}}
    }
  \pause
\\
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{\Gamma}{A}}{P}
    \jfam{{\Gamma}{B}}{Q}
    }
  { \jfameq
      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
      {\ctxwk{\ctxext{A}{P}}{Q}}
      {\ctxwk{P}{{A}{Q}}}
    }
\\
& \inference
  { \jfam{\Gamma}{A}
    \jfam{{\Gamma}{A}}{P}
    \jterm{{\Gamma}{B}}{Q}{g}
    }
  { \jtermeq
      {{{{\Gamma}{A}}{P}}{\ctxwk{P}{{A}{B}}}}
      {\ctxwk{P}{{A}{Q}}}
      {\ctxwk{\ctxext{A}{P}}{g}}
      {\ctxwk{P}{{A}{g}}}
    } 
\end{align*}
\\[\baselineskip]
\pause
To express these rules algebraically, we need to define a weakening operation on
$\famesys{\stesys}$, provided we have one on $\stesys$.
\end{frame}

\begin{frame}
\frametitle{\bf Weakening for the families algebra}
Let $\stesys$ be an extension algebra with weakening operation
$\mathbf{w}(\stesys)$. Then $\famesys{\stesys}$ has the weakening operation
$\mathbf{w}(\famesys{\stesys})$ which is uniquely determined by rendering the
diagram
\begin{small}
\begin{equation*}
\begin{tikzcd}[ampersand replacement=\&,column sep=large]
\cobesys{\stesysf_2}{\famesys{\famesys{\stesys}}}{\eft[2]}{\eft[2]}
  \ar{rr}{%
      \pullback{\beta_1}{\boldsymbol{\beta}}{\eft}{\eft}
    }
  \ar[bend right]{ddr}[swap]{\pullbackpr{1}{\eft[2]}{\eft[2]}}
  \ar[dotted]{dr}{\important{$\mathbf{w}(\famesys{\stesys})$}}
  \&
  {}\&
\cobesys{\stesysf}{\famesys{\stesys}}{\eft}{\eft}
  \ar{r}{\important{$\mathbf{w}(\stesys)$}}
  \&
\famesys{\famesys{\stesys}}
  \ar{d}{\boldsymbol{\beta}}
  \\
  {}\&
\famesys{\famesys{\famesys{\stesys}}}
  \ar{r}{\boldsymbol{\beta}_\mathbf{2}}
  \ar{d}[swap]{\eft[3]}
  \&
\famesys{\famesys{\stesys}}
  \ar{d}{\eft[2]}
  \ar{r}{\boldsymbol{\beta}}
  \&
\famesys{\stesys}
  \ar{d}{\eft}
  \\
  {}\&
\stesysf_2
  \ar{r}[swap]{\efamext}
  \&
\stesysf
  \ar{r}[swap]{\ectxext}
  \&
\stesysc
\end{tikzcd}
\end{equation*}
\end{small}
commutative.
\end{frame}
